# REPRODUCIBLE RESEARCH Project 1


# Load needed heights
```{r}
require(plyr)
require(ggplot2)

```
## loading data
```{r}

   activity <- read.csv("C:/repdata/activity.csv", header = TRUE)
   activity <- transform(activity, date= as.Date(date))

```
## WHAT IS THE MEAN

# find the total number of steps.
```{r}
     dailysteps <- ddply(activity, ~date, summarize, steps = sum(steps))

```
# plot a histogram.
```{r}
     p <- ggplot(dailysteps, aes(steps))
     p <- p + geom_histogram(fill = "red", color = "black")
     p <- p + ggtitle("Total steps for day")
     p + xlab("Steps per day")

```

# Calculate mean and mediean total number of steps taken daily.
# Average of steps taken for intervals.

```{r}
   meansteps  <- mean(dailysteps$steps, na.rm = TRUE)
   mediansteps <- median(dailysteps$steps, na.rm = TRUE)
   avgstepsperinterval <- ddply(activity, ~interval, 
                          summarize,
                          mean = mean(steps, na.rm = TRUE))

```

# 5 minute interval plot.

```{r}
   p <- ggplot(avgstepsperinterval, aes(interval, mean)) + geom_line()
   p <- p + ggtitle("Average daily activity pattern")
   p + xlab("Interval") + ylab("Number of steps")

```

# 5 min interval with max number of steps

```{r}
    maxid <- which.max(avgstepsperinterval$mean)
    maxint <- avgstepsperinterval$interval[maxid]

```

# 5 min interval ave across days.

```{r}
nasrows <-  sum(apply(is.na(activity), 1, any))

```

# Replace all NAs. 

```{r}
na.replace <- function(act) { 
     ddply(act, ~interval, function(dd) {
                steps <- dd$steps
                dd$steps[is.na(steps)] <- mean(steps, na.rm = TRUE)
                return(dd)
                })
}

```
# original dataset with no missing data.

```{r}
    activityin <- na.replace(activity)

```

# total number of steps taken by day.

```{r}
dailystepsin <- ddply(activityin, ~date, summarise, steps = sum(steps))

```
# Histogram of total number of steps.

```{r}
     p <- ggplot(dailystepsin, aes(steps))
     p <- p + geom_histogram(fill = "blue", color = "black")
     p <- p + ggtitle("Total number of steps ")
     p + xlab("Steps per day")

```

# Calculate mean and median totals.

```{r}
     dailystepsinmean <- mean(dailystepsin$steps)
     dailystepsinmedian <- median(dailystepsin$steps)
```

# Create an new weekpart weekday.

```{r}
     weekparts <- c("Weekday","Weekend")
     byday <- function(date) {
     day <- weekdays(date)
     part <- factor("Weekday", weekparts)
     
  if (day %in% c("saturday", "Sunday"))
     part <- factor("Weekend", weekparts)
     return(part)
}
activityin$weekpart <- sapply(activityin$date, byday)

```
# Panel plot containing time series of 5-min interval and average number of steps take across weekdays.

```{r}
avgsteps <- ddply(activityin, .(interval, weekpart),
                  summarise, 
                  mean = mean(steps))


        p <- ggplot(avgsteps, aes(x = interval, y = mean))
        p <- p + geom_line() + facet_grid(. ~weekpart,)
        p <- p + ggtitle("Activity on weekdays and weekends")
        p + xlab("Interval") + ylab("Number of Steps")
```
                              